### 2007-01-16 Documentation/Paper2/paper2.tex
JK - Minor typos in paper2 corrected

Neutrino oscillations are now established as the leading flavor
transition mechanism for neutrinos in a long history of many experiments, see
\eg , \Ref~\cite{Barger:2003qi}. Future facilities, using accelerator-based
 neutrino beams or a nuclear reactors as neutrino sources, are proposed
for precision measurements of the neutrino oscillation parameters. In the
simulation of these experiments, 
the presence of multiple solutions which are intrinsic to
 neutrino oscillation probabilities~\cite{Fogli:1996pv,Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
affect the performance.
Thus optimization strategies are required which maximally exploit 
complementarity between experiments. The \GLOBES\ software package~\cite{Huber:2004ka}
is a  modern experiment  simulation  and analysis tool for
a highly accurate beam and detector simulation. In addition, it provides powerful 
means to analyze correlations and degeneracies, especially for the combination
of several experiments. Compared to a Monte Carlo simulation, which yields a different
result in each run, it simulates the performance of the {\em average} experiment
(see \Ref~\cite{Schwetz:2006md} for a discussion of the meaning of ``average'').
The advantage of such an average prediction is a tremendous performance gain,
which can be used for systematical parameter space scans. In addition, it 
simplified the direct comparison of experiments.

@@

\GLOBES\ allows to simulate experiments with stationary neutrino point 
sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a continuous source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependencies can not be studied, such as  supernov\ae. 
However, in \GLOBESN\ and higher, new flexibility is introduced by the
concept of user-defined systematics. This new feature allows the cross-definition
of systematical errors over different experiments. In principle, 
this mechanism can be used for the combination of several discrete sources
and one detector, or several detectors and one source, or several sources and
several detectors. In addition, already implemented
concepts of \GLOBES\ have been used for indirect simulations of geometrical
effects. For example, the mapping of the detector location on the neutrino
energy has been simulated in \Ref~\cite{Rolinec:2006xr} by the use of
variable bin widths.

@@

The probability calculation in \GLOBES\ is based on the diagonalization of the
Hamiltonian in layers of constant matter density using the standard three flavor
scenario of neutrino oscillations. In flavor base, we have
\begin{equation}
\mathcal{H}(n_e) = \frac{1}{2E} \, U \,
\left(
\begin{array}{ccc}
0 & 0 & 0 \\
0  & \sdm & 0 \\
0  & 0 & \ldm \\
\end{array}
\right) \,
U^\dagger + 
\left(
\begin{array}{ccc}
 \pm \sqrt{2} G_F n_e & 0 & 0 \\
0 & 0 & 0 \\
0 & 0 & 0 
\end{array}
\right)
\label{equ:ham}
\end{equation}
with $U$ the mixing matrix (described by four parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, and $\deltacp$), and $n_e$ the constant electron density
in that matter density layer. The electron density is related to the matter density $\rho$ by 
$n_e \simeq 0.5 \, \rho/m_N$ with $m_N$ being the nucleon mass. The sign in the second term depends on if one
uses neutrinos or antineutrinos. For $k$ matter density layers with densities
$n^k_e$ and thicknesses $x_i$, the oscillation probability then evaluates to
\begin{equation}
P(\nu_\alpha \rightarrow \nu_\beta) = \left| \langle \nu_\beta | \mathcal{E}(n^k_e, x_k) \hdots  \mathcal{E}(n^1_e, x_1) | \nu_\alpha \rangle \right|^2
\label{equ:probs}
\end{equation}
with the evolution operators (CORRECT REFERENCE?)
\begin{equation}
  \mathcal{E}(n^k_e, x_k) = \exp \left( -i \mathcal{H}(n^k_e) x_k \right) = \frac{1}{2E} \, \hat{U}(n^k_e) \, \left( \begin{array}{ccc}
e^{-i \, m_1^k \, x_k} & 0 & 0 \\
0 &  e^{-i \, m_2^k \, x_k} & 0 \\
0 & 0 & e^{-i \, m_3^k \, x_k} 
\end{array}
\right) \, \hat{U}(n^k_e)^\dagger \, .
\end{equation}
Here the Hamiltonian $\mathcal{H}(n^k_e)$ is diagonalized by the unitary mixing matrix in matter $\hat{U}(n^k_e)$
with the eigenvalues $m_1^k$, $m_2^k$, and $m_3^k$. The probabilities in \equ{probs} are used for the
event rate computation, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
we will demonstrate how systematics is implemented thereafter.

@@

\GLOBESN\ allows the modification of this Hamiltonian and the whole probability engine without
re-compilation of the \GLOBES\ software. This is implemented
by using pointers to the functions of the standard probability engine, which can be changed if
a different probability engine is registered. For example, a user may just copy the standard probability
engine of \GLOBES , modify it, and register it. An important prerequisite for the ability to handle
more than the standard six oscillation parameters (plus matter density) $\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho)$, \ie,
\begin{equation}
\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho, \eta_1, \hdots, \eta_n)
\end{equation}
with $n$ non-standard parameters $\eta_j$. These parameters can be accessed in the usual way, see \Ref~\cite{Manual}.
For example, non-standard Hamiltonian effects in the $e$-$\tau$-sector may be introduced by modifying \equ{ham} to
\begin{equation}
 \mathcal{\hat{H}} = \mathcal{H}(n_e) + 
\left(
\begin{array}{ccc}
0 & 0 & \epsilon_{e \tau} \\
0 & 0 & 0 \\
\epsilon_{e \tau}^* & 0 & 0 
\end{array}
\right)
\label{equ:nsi}
\end{equation}
In this case, there are two more real parameters, such as the absolute value and phase of $\epsilon_{e \tau}$.
In the literature, the non-standard physics feature in \GLOBES\ as experimental feature has been used for damping effects (such as neutrino decoherence, decay, \etc) in \Ref~\cite{Blennow:2005yk}, Hamiltonian-level effects (such as non-standard matter effects) in \Ref~\cite{Blennow:2005qj}, and mass-varying neutrinos in \Ref~\cite{Schwetz:2005fy} by implementing an environment dependence of neutrino mass.  ANY OTHERS (PH!)?
We show in \figu{reactorNSI} a result from \Ref~\cite{Blennow:2005yk} provided as {\tt example6.c} in the \GLOBES\
distribution. Compared to \equ{nsi}, the underlying physics takes place at the probability level, \ie, 
\equ{probs} (for details, see \Ref~\cite{Blennow:2005yk}). In {\tt example6.c}, an analytical implementation
of this physics is used for simplicity.

@@

\GLOBES\ supports four types of systematical errors as a standard: Signal and background
normalization errors, as well as signal and background tilt or energy calibration errors.
The tilt (T) is implemented as a linear distortion of the spectrum around the center,
and the energy calibration (C) as a distortion (stretching) of the reconstructed energy scale.
For the implementation of these systematical errors, the  ``pull method'' is used~\cite{Fogli:2002pt}, which introduces  so-called nuisance parameters $\zeta_i$. 
For example, for the signal and background normalization errors, the respective event rates 
$s_i$ and $b_i$ in
each bin $i$ are multiplied with\footnote{For details on the event rate calculation, see \Ref~\cite{Huber:2004ka}.}
\begin{equation}
s_i(\zeta_1) := (1+\zeta_1)\cdot s_i \, , \quad b_i(\zeta_2) : =(1+\zeta_2)\cdot b_i \, ,
\end{equation} 
\ie, the rates are scaled by these parameters. The systematics $\chi^2$ is then minimized over
these parameters:
\begin{equation}
\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( \,
\chi^2(\boldsymbol{\lambda},
\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} \frac{\zeta_j^2}{\sigma_{\zeta_j}^2}
\, \right)\,.
\label{equ:chipull}
\end{equation}
Here $\chi^2(\boldsymbol{\lambda}, \zeta_1, \ldots, \zeta_k)$ is the usual Possonian $\chi^2$
depending on the neutrino oscillation parameters $\boldsymbol{\lambda}$ and the nuisance
parameters $\zeta_i$ (which, for instance, scale the signal or background rates).
In addition, Gaussian penalties $\zeta_j^2/\sigma_{\zeta_j}^2$ are added, where 
$\sigma_{\zeta_j}$ corresponds to the actual systematical error.
Core part of \equ{chipull} is the definition of $\chi^2$. For example, for a background-free measurement and
a signal normalization error only, it reads
\begin{equation}
  \chi^2 (\boldsymbol{\lambda},a) =  \sum_{i=1}^{\textrm{\# of bins}}2
     \left( O_{i} - (1 + a) T_{i} +
       (1 + a) T_{i} \log \frac{(1 + a) T_{i}}{O_{i}}  \right) + \frac{a^2}{\sigma_a^2} 
\label{equ:chistandard}
\end{equation}
(JK/PH: CHECK THIS EQUATION IF THIS IS THE ACTUAL IMPLEMENTATION. HERE THE SYSTEMATICS IS\\
IMPLEMENTED IN THE SIMULATED RATES. IN ADDITION, CHECK IF THEORETICAL AND EXP. RATES
ARE CORRECTLY GIVEN. [WW])

@@

In order to include correlations and degeneracies, \GLOBES\ provides
(local) marginalization routines to project the fit manifold onto
a subspace. For example, if one wants to compute the $\stheta$ precision,
one scans the $\stheta$ direction and marginalizes for each fixed $\stheta$
over the other oscillation parameters. This marginalization is performed
after the systematics $\chi^2_{\mathrm{pull}}$ in \equ{chipull} (or similar) 
has been determined:
\begin{equation}
 \chi^2_{\mathrm{proj}}(\boldsymbol{\lambda_1}) := \min_{\{ \boldsymbol{\lambda_2} \} } \left( \, \chi^2_{\mathrm{pull}} (\boldsymbol{\lambda_1} , \boldsymbol{\lambda_2} ) \, \right) \, .
\label{equ:margin}
\end{equation}
In this case, the parameter space of oscillation parameters 
$\boldsymbol{\lambda}$ is projected onto the subspace $\boldsymbol{\lambda_1}$.
If one wants to include external input which only depends on the oscillation
parameters and not on systematics, one can do this at this level. \GLOBES\ provides
the pre-defined possibility to include external Gaussian constraints on the 
oscillation parameters (and the matter density, which is treated as another
oscillation parameter). For example, an external constraint on $\theta_{12}$
is added before the marginalization in \equ{margin} by the replacement
\begin{equation}
 \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda}) \rightarrow \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda})  +
 \frac{(\theta_{12} - \theta_{12}^0)^2}{\sigma_{\theta_{12}}^2}
\label{equ:priors}
\end{equation}
with the central value $\theta_{12}^0$, where the input is added, and
the $1\sigma$ Gaussian error $\sigma_{\theta_{12}}$. This corresponds to an external
measurement of the solar mixing angle at the best-fit value $\theta_{12}^0$
with the error $\sigma_{\theta_{12}}$.

@@

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{llp{7.0cm}c}
\hline
Experiment & File name & Short description &  \Refs \\
\hline 
\multicolumn{4}{l}{\underline{Superbeam experiments:}} \\
T2K & {\tt T2K.glb} & J-PARC to Super-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
T2HK & {\tt T2HK.glb} & J-PARC to Hyper-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
NO$\nu$A & {\tt NOvA.glb} & Fermilab NuMI beamline off-axis  & \cite{Ambats:2004js,Yang_2004} \\
SPL & {\tt SPL.glb} & CERN to Fr\'{e}jus & \cite{Campagne:2006yx,Campagne:2004wt,Mezzetto:2003mm} \\[0.2cm]
\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
{\sc Reactor-I} & {\tt Reactor1.glb} & Small reactor exp., 
$\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$
& \cite{Huber:2003pm} \\
{\sc Reactor-II} & {\tt Reactor2.glb} & Large reactor exp., 
$\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ 
& \cite{Huber:2003pm} \\
Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & Double Chooz near 
detector & \cite{Huber:2006vr} \\ 
 & {\tt D-Chooz\_far.glb} & Double Chooz far detector & \\[0.2cm]
\multicolumn{4}{l}{\underline{Beta beams:}} \\
Low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus 
scenario & \cite{Campagne:2006yx} \\
Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ green field scenario 
& \cite{Burguet-Castell:2005pa} \\
Variable $\gamma$ & {\tt BBvar\_WC.glb} & Variable $\gamma$ beta beam with water Cherenkov detector, $50 \lesssim \gamma \lesssim 500$ & \cite{Huber:2005jk} \\[0.2cm]
	& {\tt BBvar\_TASD.glb} & Variable $\gamma$ beta beam with totally active scint. detector (TASD), 
$100 \lesssim \gamma \lesssim 3\, 000$ & \cite{Huber:2005jk} \\[0.4cm]
\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
Standard & {\tt NFstandard.glb} & Standard neutrino factory, $50 \, \mathrm{kt}$ magnetized iron calorimeter, 
$E_\mu = 50 \, \mathrm{GeV}$ & \cite{Huber:2002mx} \\
Variable $E_\mu$ & {\tt NFvar.glb} & Variable neutrino factory, disapp. channels
{\em without} CID; $10 \, \mathrm{GeV} \lesssim E_\mu \lesssim 80 \, \mathrm{GeV}$ & \cite{Huber:2002mx,Huber:2006wb}  \\
Gold + Silver & {\tt NF\_GoldSilver.glb} & As {\tt NFvar.glb} 
plus 5~kt ECC detector for Silver Channel
measurement & \cite{Huber:2002mx,Huber:2006wb,Autiero:2003fu} \\
Hybrid detector & {\tt NF\_hR\_lT.glb} & As {\tt NFvar.glb}, but 
lower threshold and better energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\
\hline
\end{tabular}
\end{center}
} % small
\mycaption{\label{tab:experiments} 
Pre-defined experiment prototypes coming with \GLOBESN : The columns represent the \AEDL\ filename, 
a short description, and the references in which they are originally used and 
discussed (except from minor modifications, such as a different implementation of 
the energy threshold function). 
%
IS THE GAMMA=350 OPTION REALLY GREEN-FIELD? I THOUGHT THAT THIS MAY BE THE MAXIMUM
ONE COULD GET OUT OF THE SPS WITH ACCELERATOR MODIFICATIONS? [WW]
}
\end{table}

\AEDL\ (Abstract Experiment Definition Language) is a special syntax developed for the description of
experiments in \GLOBES\ represented by simple text files. Core part of \AEDL\ are the following constructions
(for details, see \Ref~\cite{Manual}):
\begin{description}
\item[Channel] A channel represents an oscillation channel from the source flux, over the initial and final
flavors and the use of neutrinos or antineutrinos, to the interaction type and cross section. The result
is an event rate vector.
\item[Rule] A rule combines one or more signal channels with one or more background channels. The rates
from these channels are added, and the signal and background are implemented with specific systematics, \ie,
systematics is connected to the concept of a rule.
The result of a rule is a systematics $\chi^2$ corresponding to \equ{chipull}.
\item[Experiment] An experiment is a combination of one or more rules, which may correspond to different
appearance and disappearance channels, neutrino and antineutrino operations modes, \etc. An experiment has a common detector and baseline, but may use more different fluxes (such as for neutrinos and antineutrinos). Therefore,
the matter density profile is the same for all rules of an experiment. The result of an experiment
is again a systematics $\chi^2$, but summed over all rules.
\item[Experiment combination] An experiment combination consists of different experiments with different detectors, baselines, and (uncorrelated) matter density profiles. The result of an experiment combination
is again a systematics $\chi^2$, but summed over all experiments.
\end{description}
Note that after the $\chi^2$ is summed over all rules or experiments as specified by the user in the application software, \equ{margin} is applied, and the oscillation parameters are marginalized over.

In \GLOBESN , this construction of experiments is extended by user-defined systematics, which allows
the user to break up the strict coupling between systematics and rules. For example, systematics can be
correlated between different rules or experiments, such as for an reactor experiment with identical
near and far detectors. In addition, one can introduce user-defined systematical errors.
For that purpose, the former error dimension concept has been changed and
extended. For example, one has to give each user-defined systematics a name, and this name has
to be matched by the application software. 
%
Further changes in \AEDL\ are the requirement to specify the minimum \GLOBES\ version the \AEDL\ is running with,
and the extensions by lists as \AEDL\ variables and by an interpolation feature.

>>>>

Neutrino oscillations are now established as the leading flavor
transition mechanism for neutrinos in a long history of many experiments, see
\eg , \Ref~\cite{Barger:2003qi}. Future facilities, using accelerator-based
 neutrino beams or nuclear reactors as neutrino sources, are proposed
for precision measurements of the neutrino oscillation parameters. In the
simulation of these experiments, 
the presence of multiple solutions which are intrinsic to
 neutrino oscillation probabilities~\cite{Fogli:1996pv,Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
affect the performance.
Thus, optimization strategies are required which maximally exploit 
complementarity between experiments. The \GLOBES\ software package~\cite{Huber:2004ka}
is a  modern experiment  simulation  and analysis tool for
a highly accurate beam and detector simulation. In addition, it provides powerful 
means to analyze correlations and degeneracies, especially for the combination
of several experiments. Compared to a Monte Carlo simulation, which yields a different
result in each run, it simulates the performance of the {\em average} experiment
(see \Ref~\cite{Schwetz:2006md} for a discussion of the meaning of ``average'').
The advantage of such an average prediction is a tremendous performance gain,
which can be used for systematical parameter space scans. In addition, it 
simplifies the direct comparison of experiments.

@@

\GLOBES\ allows to simulate experiments with stationary neutrino point 
sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a continuous source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependence can not be studied, such as  supernov\ae. 
However, in \GLOBESN\ and higher, new flexibility is introduced by the
concept of user-defined systematics. This new feature allows the cross-definition
of systematical errors over different experiments. In principle, 
this mechanism can be used for the combination of several discrete sources
and one detector, or several detectors and one source, or several sources and
several detectors. In addition, already implemented
concepts of \GLOBES\ have been used for indirect simulations of geometrical
effects. For example, the mapping of the detector location on the neutrino
energy has been simulated in \Ref~\cite{Rolinec:2006xr} by the use of
variable bin widths.

@@

The probability calculation in \GLOBES\ is based on the diagonalization of the
Hamiltonian in layers of constant matter density using the standard three flavor
scenario of neutrino oscillations. In the flavor basis, we have
\begin{equation}
\mathcal{H}(n_e) = \frac{1}{2E} \, U \,
\left(
\begin{array}{ccc}
0 & 0 & 0 \\
0  & \sdm & 0 \\
0  & 0 & \ldm \\
\end{array}
\right) \,
U^\dagger + 
\left(
\begin{array}{ccc}
 \pm \sqrt{2} G_F n_e & 0 & 0 \\
0 & 0 & 0 \\
0 & 0 & 0 
\end{array}
\right)
\label{equ:ham}
\end{equation}
with $U$ being the mixing matrix (described by four parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, and $\deltacp$), and $n_e$ the constant electron density
in the respective matter density layer. The electron density is related to the matter density $\rho$ by 
$n_e \simeq 0.5 \, \rho/m_N$ with $m_N$ being the nucleon mass. The sign in the second term depends on whether one
uses neutrinos or antineutrinos. For $k$ matter density layers with densities
$n^k_e$ and thicknesses $x_i$, the oscillation probability then evaluates to
\begin{equation}
P(\nu_\alpha \rightarrow \nu_\beta) = \left| \langle \nu_\beta | \mathcal{E}(n^k_e, x_k) \hdots  \mathcal{E}(n^1_e, x_1) | \nu_\alpha \rangle \right|^2
\label{equ:probs}
\end{equation}
with the evolution operators (CORRECT REFERENCE?)
\begin{equation}
  \mathcal{E}(n^k_e, x_k) = \exp \left( -i \mathcal{H}(n^k_e) x_k \right) = \frac{1}{2E} \, \hat{U}(n^k_e) \, \left( \begin{array}{ccc}
e^{-i \, m_1^k \, x_k} & 0 & 0 \\
0 &  e^{-i \, m_2^k \, x_k} & 0 \\
0 & 0 & e^{-i \, m_3^k \, x_k} 
\end{array}
\right) \, \hat{U}(n^k_e)^\dagger \, .
\end{equation}
Here the Hamiltonian $\mathcal{H}(n^k_e)$ is diagonalized by the unitary mixing matrix in matter $\hat{U}(n^k_e)$
with the eigenvalues $m_1^k$, $m_2^k$, and $m_3^k$. The probabilities in \equ{probs} are used for the
event rate computation, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
we will demonstrate how systematics is implemented thereafter.

@@

\GLOBESN\ allows the modification of this Hamiltonian and the whole probability engine without
re-compilation of the \GLOBES\ software. This is implemented
by using pointers to the functions of the standard probability engine, which can be changed by
registering a different probability engine. For example, a user may just copy the standard probability
engine of \GLOBES , modify it, and register it. An important prerequisite for the ability to handle
more than the standard six oscillation parameters (plus matter density) $\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho)$, \ie,
\begin{equation}
\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho, \eta_1, \hdots, \eta_n)
\end{equation}
with $n$ non-standard parameters $\eta_j$. These parameters can be accessed in the usual way, see \Ref~\cite{Manual}.
For example, non-standard Hamiltonian effects in the $e$-$\tau$-sector may be introduced by modifying \equ{ham} to
\begin{equation}
 \mathcal{\hat{H}} = \mathcal{H}(n_e) + 
\left(
\begin{array}{ccc}
0 & 0 & \epsilon_{e \tau} \\
0 & 0 & 0 \\
\epsilon_{e \tau}^* & 0 & 0 
\end{array}
\right)
\label{equ:nsi}
\end{equation}
In this case, there are two more real parameters, such as the absolute value and phase of $\epsilon_{e \tau}$.
In the literature, the non-standard physics feature in \GLOBES\ as experimental feature has been used for damping effects (such as neutrino decoherence, decay, \etc) in \Ref~\cite{Blennow:2005yk}, Hamiltonian-level effects (such as non-standard matter effects) in \Ref~\cite{Blennow:2005qj}, and mass-varying neutrinos in \Ref~\cite{Schwetz:2005fy} by implementing an environment dependence of neutrino mass.  ANY OTHERS (PH!)?
We show in \figu{reactorNSI} a result from \Ref~\cite{Blennow:2005yk} provided as {\tt example6.c} in the \GLOBES\
distribution. Compared to \equ{nsi}, the underlying physics takes place at the probability level, \ie, 
\equ{probs} (for details, see \Ref~\cite{Blennow:2005yk}). In {\tt example6.c}, an analytical implementation
of this physics is used for simplicity.

@@

\GLOBES\ supports four types of systematical errors by default: Signal and background
normalization errors, as well as signal and background tilt or energy calibration errors.
The tilt (T) is implemented as a linear distortion of the spectrum around the center,
and the energy calibration (C) as a distortion (stretching) of the reconstructed energy scale.
For the implementation of these systematical errors, the  ``pull method'' is used~\cite{Fogli:2002pt}, which introduces  so-called nuisance parameters $\zeta_i$. 
For example, for the signal and background normalization errors, the respective event rates 
$s_i$ and $b_i$ in
each bin $i$ are multiplied with\footnote{For details on the event rate calculation, see \Ref~\cite{Huber:2004ka}.}
\begin{equation}
s_i(\zeta_1) := (1+\zeta_1)\cdot s_i \, , \quad b_i(\zeta_2) : =(1+\zeta_2)\cdot b_i \, ,
\end{equation} 
\ie, the rates are scaled by these parameters. The systematics $\chi^2$ is then minimized over
these parameters:
\begin{equation}
\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( \,
\chi^2(\boldsymbol{\lambda},
\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} \frac{\zeta_j^2}{\sigma_{\zeta_j}^2}
\, \right)\,.
\label{equ:chipull}
\end{equation}
Here $\chi^2(\boldsymbol{\lambda}, \zeta_1, \ldots, \zeta_k)$ is the usual Possonian $\chi^2$
depending on the neutrino oscillation parameters $\boldsymbol{\lambda}$ and the nuisance
parameters $\zeta_i$ (which, for instance, scale the signal or background rates).
In addition, Gaussian penalties $\zeta_j^2/\sigma_{\zeta_j}^2$ are added, where 
$\sigma_{\zeta_j}$ corresponds to the actual systematical error.
The core part of \equ{chipull} is the definition of $\chi^2$. For example, for a background-free measurement and
a signal normalization error only, it reads
\begin{equation}
  \chi^2 (\boldsymbol{\lambda},a) =  \sum_{i=1}^{\textrm{\# of bins}}2
     \left( O_{i} - (1 + a) T_{i} +
       (1 + a) T_{i} \log \frac{(1 + a) T_{i}}{O_{i}}  \right) + \frac{a^2}{\sigma_a^2} 
\label{equ:chistandard}
\end{equation}
(JK/PH: CHECK THIS EQUATION IF THIS IS THE ACTUAL IMPLEMENTATION. HERE THE SYSTEMATICS IS\\
IMPLEMENTED IN THE SIMULATED RATES. IN ADDITION, CHECK IF THEORETICAL AND EXP. RATES
ARE CORRECTLY GIVEN. [WW])

@@

In order to include correlations and degeneracies, \GLOBES\ provides
(local) marginalization routines to project the fit manifold onto
a subspace. For example, if one wants to compute the $\stheta$ precision,
one scans the $\stheta$ direction and marginalizes for each fixed $\stheta$
over the other oscillation parameters. This marginalization is performed
after the systematics $\chi^2_{\mathrm{pull}}$ in \equ{chipull} (or similar) 
has been determined:
\begin{equation}
 \chi^2_{\mathrm{proj}}(\boldsymbol{\lambda_1}) := \min_{\{ \boldsymbol{\lambda_2} \} } \left( \, \chi^2_{\mathrm{pull}} (\boldsymbol{\lambda_1} , \boldsymbol{\lambda_2} ) \, \right) \, .
\label{equ:margin}
\end{equation}
In this case, the parameter space of oscillation parameters 
$\boldsymbol{\lambda}$ is projected onto the subspace $\boldsymbol{\lambda_1}$.
If one wants to include external input, which only depends on the oscillation
parameters and not on systematics, one can do so at this level. \GLOBES\ provides
the pre-defined possibility to include external Gaussian constraints on the 
oscillation parameters (and the matter density, which is treated as another
oscillation parameter). For example, an external constraint on $\theta_{12}$
is added before the marginalization in \equ{margin} by the replacement
\begin{equation}
 \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda}) \rightarrow \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda})  +
 \frac{(\theta_{12} - \theta_{12}^0)^2}{\sigma_{\theta_{12}}^2}
\label{equ:priors}
\end{equation}
with the central value $\theta_{12}^0$, where the input is added, and
the $1\sigma$ Gaussian error $\sigma_{\theta_{12}}$. This corresponds to an external
measurement of the solar mixing angle at the best-fit value $\theta_{12}^0$
with the error $\sigma_{\theta_{12}}$.

@@

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{llp{7.0cm}c}
\hline
Experiment & File name & Short description &  \Refs \\
\hline 
\multicolumn{4}{l}{\underline{Superbeam experiments:}} \\
T2K & {\tt T2K.glb} & J-PARC to Super-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
T2HK & {\tt T2HK.glb} & J-PARC to Hyper-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
NO$\nu$A & {\tt NOvA.glb} & Fermilab NuMI beamline off-axis  & \cite{Ambats:2004js,Yang_2004} \\
SPL & {\tt SPL.glb} & CERN to Fr\'{e}jus & \cite{Campagne:2006yx,Campagne:2004wt,Mezzetto:2003mm} \\[0.2cm]
\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
{\sc Reactor-I} & {\tt Reactor1.glb} & Small reactor exp., 
$\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$
& \cite{Huber:2003pm} \\
{\sc Reactor-II} & {\tt Reactor2.glb} & Large reactor exp., 
$\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ 
& \cite{Huber:2003pm} \\
Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & Double Chooz near 
detector & \cite{Huber:2006vr} \\ 
 & {\tt D-Chooz\_far.glb} & Double Chooz far detector & \\[0.2cm]
\multicolumn{4}{l}{\underline{Beta beams:}} \\
Low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus 
scenario & \cite{Campagne:2006yx} \\
Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ green field scenario 
& \cite{Burguet-Castell:2005pa} \\
Variable $\gamma$ & {\tt BBvar\_WC.glb} & Variable $\gamma$ beta beam with water Cherenkov detector, $50 \lesssim \gamma \lesssim 500$ & \cite{Huber:2005jk} \\[0.2cm]
	& {\tt BBvar\_TASD.glb} & Variable $\gamma$ beta beam with totally active scint. detector (TASD), 
$100 \lesssim \gamma \lesssim 3\, 000$ & \cite{Huber:2005jk} \\[0.4cm]
\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
Standard & {\tt NFstandard.glb} & Standard neutrino factory, $50 \, \mathrm{kt}$ magnetized iron calorimeter, 
$E_\mu = 50 \, \mathrm{GeV}$ & \cite{Huber:2002mx} \\
Variable $E_\mu$ & {\tt NFvar.glb} & Variable neutrino factory, disapp. channels
{\em without} CID; $10 \, \mathrm{GeV} \lesssim E_\mu \lesssim 80 \, \mathrm{GeV}$ & \cite{Huber:2002mx,Huber:2006wb}  \\
Gold + Silver & {\tt NF\_GoldSilver.glb} & As {\tt NFvar.glb} 
plus 5~kt ECC detector for Silver Channel
measurement & \cite{Huber:2002mx,Huber:2006wb,Autiero:2003fu} \\
Hybrid detector & {\tt NF\_hR\_lT.glb} & As {\tt NFvar.glb}, but 
lower threshold and better energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\
\hline
\end{tabular}
\end{center}
} % small
\mycaption{\label{tab:experiments} 
Pre-defined experiment prototypes coming with \GLOBESN : The columns represent the \AEDL\ filename, 
a short description, and the references in which the files are originally used and 
discussed (except for minor modifications, such as a different implementation of 
the energy threshold function). 
%
IS THE GAMMA=350 OPTION REALLY GREEN-FIELD? I THOUGHT THAT THIS MAY BE THE MAXIMUM
ONE COULD GET OUT OF THE SPS WITH ACCELERATOR MODIFICATIONS? [WW]
}
\end{table}

\AEDL\ (Abstract Experiment Definition Language) is a special syntax developed for the description of
experiments in \GLOBES, represented by simple text files. Core part of \AEDL\ are the following constructions
(for details, see \Ref~\cite{Manual}):
\begin{description}
\item[Channel] A channel represents an oscillation channel from the source flux, over the initial and final
flavors and the use of neutrinos or antineutrinos, to the interaction type and cross section. The result
is an event rate vector.
\item[Rule] A rule combines one or more signal channels with one or more background channels. The rates
from these channels are added, and the signal and background are implemented with specific systematics, \ie,
systematics is connected to the concept of a rule.
The result of a rule is a systematics $\chi^2$ corresponding to \equ{chipull}.
\item[Experiment] An experiment is a combination of one or more rules, which may correspond to different
appearance and disappearance channels, neutrino and antineutrino operations modes, \etc. An experiment has a common detector and baseline, but may use several different fluxes (such as for neutrinos and antineutrinos). Therefore,
the matter density profile is the same for all rules of an experiment. The result of an experiment
is again a systematics $\chi^2$, but summed over all rules.
\item[Experiment combination] An experiment combination consists of different experiments with different detectors, baselines, and (uncorrelated) matter density profiles. The result of an experiment combination
is again a systematics $\chi^2$, but summed over all experiments.
\end{description}
Note that after the $\chi^2$ is summed over all rules or experiments as specified by the user in the application software, \equ{margin} is applied, and the oscillation parameters are marginalized over.

In \GLOBESN , this construction of experiments is extended by user-defined systematics, which allows
the user to break up the strict coupling between systematics and rules. For example, systematics can be
correlated between different rules or experiments, such as for a reactor experiment with identical
near and far detectors. In addition, one can introduce user-defined systematical errors.
For that purpose, the former error dimension concept has been changed and
extended. For example, one has to give each user-defined systematics a name, and this name has
to be matched by the application software. 
%
Further changes in \AEDL\ are the requirement to specify the minimum \GLOBES\ version the \AEDL\ file can be used with,
and the extensions by lists as \AEDL\ variables and by an interpolation feature.
